\bf Estimating the number of clusters using Cross Validation

### 2016-01-28 New simulation/technometrics.tex
Update real data example--high dimension data

\section{Real data application}
We also applied our proposed method to three real world data sets obtained
from the University of California Irvine machine learning repository. The
first and third data sets are selected because there are clear number of
clusters in those two data sets. The second data set is used as a benchmark
data set since it was widely used in literature.

@@

The third data set is the Sonar data returned from two targets -- a metal
cylinder and a rock with similar shape, which is first studied by
\cite{gorman1988analysis} using a neural network. Both targets were impinged
by pulse which was a wide-band linear FM chirp $(ka = 55.6)$. Returns were
collected at a range of $10$ meters and obtained from the cylinder at aspect
angles spanning $90^{\circ}$ and from the rock at aspect angles spanning
$180^{\circ}$. The data set contains $208$ returns ($111$ cylinder returns and
$97$ rock returns), with each composed of $60$ spectral samples, normalized to
take on values between $0$ and $1$. So the data has $60$ features with clearly
$2$ clusters.

\begin{table}[H]
\begin{center}
\captionsetup{justification=centering}
\caption{\label{table2} Number of clusters selected by each algorithm}
\begin{tabular}{lccccc}
    \hline                  
 & Congress Voting && Breast Cancer && Sonar \\ \hline                    
CH-index & $2$ && $2$  && $2$    \\
Hartigan & $3$ && $3$  &&  $3$   \\
Jump & $9$ && $9$ && $10$  \\   
Prediction strength & $2$ &&  $2$ && $1$    \\
Bootstrap stability & $2$ &&  $2$ &&  $10$   \\ 
Gap & $10$ && $9$ && $10$  \\   
Gaussian-Mix & $7$ && $5$ && $1$  \\   
Gabriel & $2$  && $3$  && $2$  \\    
Wold & $2$ && $3$ && $10$ \\  \hline 
\end{tabular}
\end{center}
\hspace{0.5in} \footnotesize {All the algorithms executed with their default parameter settings with $k$ ranges from $1$ to $10$}
\end{table}

Since most congressmen vote base on their parties' interest, $2$ parties
(Democratic and Republican) represent two clusters in this data set. So the
optimal number should be two. Close inspection shows $k$-means with
$k=2$ separates the two parties very well with the lowest miss-classification
error $10.43\%$ ($k=3$ has $14.78\%$). Note that CH-index and Bootstrap
stability also return $k=2$, but $2$ is the lower bound those methods can
select for $k$. So it's not clear they actually choose $k=2$ or they hit the
lower bound (they would pick $k=1$ if allowed). For the breast cancer data,
it's known to have at least $2$ cluster based on whether it's benign specimens
or not. But it doesn't mean the optimal $k$ should be $2$.
\cite{fujita2014non} noticed that the malign group is quite heterogeneous and
can be further clustered into at least two subgroups. Hence, the result $k=3$
given by our proposed Gabriel method (as well as Wold method) make sense. For
the Sonar data set which is relatively high-dimension, only our proposed
Gabriel method and CH-index selects $k=2$. Majority of the methods collapse to
either select the maximum or select the minimum number, underline how
difficult it was to pick the right $k$ when dimension increases.

>>>>

\section{Real data application}
We also applied our proposed method to three real world data sets, two were obtained
from the University of California Irvine machine learning repository. The
first and third data sets are selected because there are clear number of
clusters in those two data sets. The second data set is used as a benchmark
data set since it was widely used in literature.

@@

The third data set is gene expression data of $5$ types of brain tumours from \cite{de2008clustering}, which contains $42$ observations including $10$ medulloblastomas, $10$ malignant gliomas, $10$ atypical teratoid/rhabdoid tumours (AT/RTs), $8$ primitive neuroectodermal tumours (PNETs) and $4$ normal cerebella. It was originally used by \cite{pomeroy2002prediction} to show that medulloblastomas are molecularly distinct from other types of brain tumours. There were $7129$ genes originally, but after feature selection, $1379$ genes left. So the data set is $42 \times 1379$, which serves as a high dimensional example.

\begin{table}[H]
\begin{center}
\captionsetup{justification=centering}
\caption{\label{table2} Number of clusters selected by each algorithm}
\begin{tabular}{lccccc}
    \hline                  
 & Congress Voting && Breast Cancer && Brain Tumours \\ \hline                    
CH-index & $2$ && $2$  && $2$    \\
Hartigan & $3$ && $3$  &&  $4$   \\
Jump & $10$ && $9$ && $1$  \\   
Prediction strength & $2$ &&  $2$ && $1$    \\
Bootstrap stability & $2$ &&  $2$ &&  $7$   \\ 
Gap & $8$ && $10$ && $10$  \\   
Gaussian-Mix & $2$ && $5$ && $2$  \\   
Gabriel & $2$  && $3$  && $5$  \\    
Wold & $2$ && $3$ && $4$ \\  \hline

\end{tabular}
\end{center}
\hspace{0.5in} \footnotesize {All the algorithms executed with their default parameter settings with $k$ ranges from $1$ to $10$}
\end{table}

Since most congressmen vote base on their parties' interest, $2$ parties
(Democratic and Republican) represent two clusters in this data set. So the
optimal number should be two. Close inspection shows $k$-means with
$k=2$ separates the two parties very well with the lowest miss-classification
error $10.43\%$ ($k=3$ has $14.78\%$). Note that CH-index and Bootstrap
stability also return $k=2$, but $2$ is the lower bound those methods can
select for $k$. So it's not clear they actually choose $k=2$ or they hit the
lower bound (they would pick $k=1$ if allowed). For the breast cancer data,
it's known to have at least $2$ cluster based on whether it's benign specimens
or not. But it doesn't mean the optimal $k$ should be $2$.
\cite{fujita2014non} noticed that the malign group is quite heterogeneous and
can be further clustered into at least two subgroups. Hence, the result $k=3$
given by our proposed Gabriel method make sense. For
the brain tumours gene expression data set which is high-dimension, it is also clear that 
the correct number of clusters should be $5$, each cluster corresponding to one type of 
brain tumour. From the table we can see only our proposed
Gabriel method correctly selects $k=5$, underline how
difficult it was to pick the right $k$ when dimension is high.
